---
title: "Exercise 5"
format: docx
---


Customers are often analyzed based on those who have made recent purchases, frequent purchases, and spent the most. This is known as a recency, frequency, monetary (RFM) analysis. Now that we can tidy data, we can analyze customers based on recent purchases and frequent purchases. In RStudio, create a new Quarto document and do the following.

1. How would you define customers who have made "recent" purchases? State how you are defining recency, and use R to report on the composition of these customers.
2. How would you define customers who have made "frequent" purchases? State how you are defining frequency, and use R to report on the composition of these customers.
3. Render the Quarto document into Word and upload to Canvas.

**Five points total, one point each for:**

- **Tidy and transform the data to identify recent customers.**
- **Explore the composition of recent customers and provide an explanation.**
- **Tidy and transform the data to identify frequent customers.**
- **Explore the composition of frequent customers and provide an explanation.**
- **Submitting a rendered Word document.**

## Recency

For one possible definition of recency, let's determine which customers have made more than 10 transactions in 2018.

```{r}
# Load packages.
library(tidyverse)
```

```{r echo=FALSE}
# Import data.
crm_data <- read_csv("customer_data.csv", show_col_types = F) |> 
  left_join(read_csv("store_transactions.csv", show_col_types = F), join_by(customer_id))
```

```{r}
# Select transaction data for 2018, tidy, summarize purchase history,
# and filter on respondents who have made more than 10 transactions.
recent_cust <- crm_data |> 
  select(customer_id, contains("2018")) |> 
  pivot_longer(
    -customer_id,
    names_to = "month_year",
    values_to = "transactions"
  ) |> 
  group_by(customer_id) |> 
  summarize(total_transactions = sum(transactions)) |> 
  filter(total_transactions > 10)
  
recent_cust
```

Now that we have identified which customers satisfy this definition of "recent customers," we can use a `left_join()` to join the remainder of the `crm_data` and explore these customers' composition.

```{r}
# Join crm_data with recent_cust.
recent_cust <- recent_cust |> 
  left_join(crm_data, join_by(customer_id))

# Explore distributions of continuous variables.
recent_cust |> 
  mutate(age = 2026 - birth_year) |> 
  select(total_transactions, income, credit, age) |> 
  pivot_longer(
    total_transactions:age,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_histogram() +
  facet_wrap(~ var_names, scales = "free") +
  labs(x = "Continuous Variables",
       y = "Count",
       title = "Continuous Variable Distributions for Recent Customers")
```

Recent customers tend to be below 60 years old. Credit scores vary substantially, with most around 650. Most incomes are below 200,000, with a few in the upper tail that approach 400,000. In terms of `total_transactions`, we see a long tail above 10 transactions, up to 35.

```{r}
# Explore distributions of discrete variables.
recent_cust |> 
  select(gender, married, college_degree, region) |> 
  pivot_longer(
    gender:region,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_bar() +
  facet_wrap(~ var_names, scales = "free") +
  labs(x = "Categorical Variables",
       y = "Count",
       title = "Categorical Variable Distributions for Recent Customers")
```

Most recent customers have a college degree, are female, are unmarried, and tend to shop in the Northeast and West regions.

## Frequency

For one possible definition of frequency, let's determine which customers have made more than 150 transactions overall.

```{r}
# Tidy and summarize purchase history and filter on respondents
# who have made more than 150 transactions overall.
frequent_cust <- crm_data |> 
  select(customer_id, jan_2005:dec_2018) |> 
  pivot_longer(
    -customer_id,
    names_to = "month_year",
    values_to = "transactions"
  ) |> 
  group_by(customer_id) |> 
  summarize(total_transactions = sum(transactions)) |> 
  filter(total_transactions > 150)
  
frequent_cust
```

Now that we have identified which customers satisfy this definition of "frequent customers," we can use a `left_join()` to join the remainder of the `crm_data` and explore these customers' composition.

```{r}
# Join crm_data with frequent_cust.
frequent_cust <- frequent_cust |> 
  left_join(crm_data, join_by(customer_id))

# Explore distributions of continuous variables.
frequent_cust |> 
  mutate(age = 2026 - birth_year) |> 
  select(total_transactions, income, credit, age) |> 
  pivot_longer(
    total_transactions:age,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_histogram() +
  facet_wrap(~ var_names, scales = "free") +
  labs(x = "Continuous Variables",
       y = "Count",
       title = "Continuous Variable Distributions for Frequent Customers")
```

Frequent customers look fairly similar to the recent customers on continuous variables, with perhaps a slightly higher average credit score.

```{r}
# Explore distributions of discrete variables.
frequent_cust |> 
  select(gender, married, college_degree, region) |> 
  pivot_longer(
    gender:region,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_bar() +
  facet_wrap(~ var_names, scales = "free") +
  labs(x = "Categorical Variables",
       y = "Count",
       title = "Categorical Variable Distributions for Frequent Customers")
```

Similarly, most frequent customers have a college degree, are female, are unmarried, and tend to shop in the Northeast and West regions.

