---
title: "Exercise 5"
format: docx
---


Customers are often analyzed based on those who have made recent purchases, frequent purchases, and spent the most. This is known as a recency, frequency, monetary (RFM) analysis. Now that we can tidy data, we can analyze customers based on recent purchases and frequent purchases. In RStudio, create a new Quarto document and do the following.

1. How would you define customers who have made "recent" purchases? Remember, the most recent data we have for this exercise is from 2018. State how you are defining recency, and use R to report on the composition of these customers.
2. How would you define customers who have made "frequent" purchases? Remember, the most recent data we have for this exercise is from 2018. State how you are defining frequency, and use R to report on the composition of these customers.
3. Render the Quarto document into Word and upload to Canvas.

**Five points total, one point each for:**

- **Recode the data to identify recent customers.**
- **Explore the composition of recent customers and provide an explanation.**
- **Recode the data to identify frequent customers.**
- **Explore the composition of frequent customers and provide an explanation.**
- **Submitting a rendered Word document.**

## Recency

For one possible definition of recency, let's determine which customers have made more than 10 transactions in 2018.

```{r}
# Load packages.
library(tidyverse)
```

```{r eval=FALSE}
# Import data.
crm_data <- read_csv("customer_data.csv", show_col_types = F) |> 
  left_join(read_csv("store_transactions.csv", show_col_types = F), join_by(customer_id))
```

```{r echo=FALSE}
# Import data.
crm_data <- read_csv("customer_data.csv", show_col_types = F) |> 
  left_join(read_csv("store_transactions.csv", show_col_types = F), join_by(customer_id))
```

```{r}
# Select transaction data for 2018, tidy, summarize purchase history,
# and filter on respondents who have made more than 10 transactions.
recent_cust <- crm_data |> 
  select(customer_id, contains("2018")) |> 
  pivot_longer(
    -customer_id,
    names_to = "month_year",
    values_to = "transactions"
  ) |> 
  group_by(customer_id) |> 
  summarize(total_transactions = sum(transactions)) |> 
  filter(total_transactions > 10)
  
recent_cust
```

Now that we have identified which customers satisfy this definition of "recent customers," we can use a `left_join()` to join the remainder of the `crm_data` and explore these customers' composition.

```{r}
# Join crm_data with recent_cust.
recent_cust <- recent_cust |> 
  left_join(crm_data, join_by(customer_id))

# Explore distributions of continuous variables.
recent_cust |> 
  mutate(age = 2021 - birth_year) |> 
  select(total_transactions, income, credit, age) |> 
  pivot_longer(
    total_transactions:age,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_histogram() +
  facet_wrap(~ var_names, scales = "free")
```

Recent customers don't look so dissimilar from the overall customer base in terms of continuous variables: same sort of `age`, `credit`, and `income` distributions. In terms of `total_transactions`, we see a long tail above 10 transactions, up to 35.

```{r}
# Explore distributions of discrete variables.
recent_cust |> 
  select(gender, married, college_degree, region) |> 
  pivot_longer(
    gender:region,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_bar() +
  facet_wrap(~ var_names, scales = "free")
```

Most recent customers have a college degree, are female, and are unmarried. They appear to be fairly evenly distributed across regions, compared to the overall customer base.

## Frequency

For one possible definition of frequency, let's determine which customers have made more than 150 transactions overall.

```{r}
# Tidy and summarize purchase history and filter on respondents
# who have made more than 150 transactions overall.
frequent_cust <- crm_data |> 
  select(customer_id, jan_2005:dec_2018) |> 
  pivot_longer(
    -customer_id,
    names_to = "month_year",
    values_to = "transactions"
  ) |> 
  group_by(customer_id) |> 
  summarize(total_transactions = sum(transactions)) |> 
  filter(total_transactions > 150)
  
frequent_cust
```

Now that we have identified which customers satisfy this definition of "frequent customers," we can use a `left_join()` to join the remainder of the `crm_data` and explore these customers' composition.

```{r}
# Join crm_data with frequent_cust.
frequent_cust <- frequent_cust |> 
  left_join(crm_data, join_by(customer_id))

# Explore distributions of continuous variables.
frequent_cust |> 
  mutate(age = 2021 - birth_year) |> 
  select(total_transactions, income, credit, age) |> 
  pivot_longer(
    total_transactions:age,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_histogram() +
  facet_wrap(~ var_names, scales = "free")
```

Frequent customers don't look so dissimilar from the overall customer base in terms of continuous variables: maybe skewed slightly younger but `credit` and `income` distributed the same. In terms of `total_transactions`, we see a long tail above 150 transactions, up to nearly 250 transactions total.

```{r}
# Explore distributions of discrete variables.
frequent_cust |> 
  select(gender, married, college_degree, region) |> 
  pivot_longer(
    gender:region,
    names_to = "var_names",
    values_to = "var_values"
  ) |> 
  ggplot(aes(x = var_values)) +
  geom_bar() +
  facet_wrap(~ var_names, scales = "free")
```

Similarly, most frequent customers have a college degree, are female, and are unmarried. They appear to be fairly evenly distributed across regions, compared to the overall customer base.

